### 2025-09-25 paper/sec/evaluation.tex
liakata12.

RQ1. How robust is the system when provided with incomplete or ambiguous information?
In real-world authoring tasks, text fragments may omit key criteria (e.g. \textbf{?}) or contain vague identifiers (e.g. \textbf{?} without further context).
We therefore ask how performance degrades when the specification is underspecified or when multiple interpretations are possible. Here, we distinguish between two forms of ambiguity:
(1) domain-level ambiguity, which reflects genuine uncertainty in the data (e.g. two models achieving identical scores).
(2) specification-level ambiguity, which arises from incomplete or imprecise descriptions in the query (e.g. missing fields that would otherwise resolve the expression uniquely).

RQ2. How sensitive is the system to misleading or adversarial information?
Authors may, intentionally or unintentionally, introduce misleading elements into the text. These include inconsistent identifiers, incorrect numerical values, or contradictory claims.
We ask how the system responds in such cases: does it fail gracefully, does it default to literal reproduction of the misleading fragment, or does it attempt to generate a plausible, but incorrect, expression?

Together, these questions probe the limits of the InterpretationAgent's ability to generate reliable code. By contrasting genuine uncertainty with misleading or ill-specified inputs, we aim to understand not only the model's raw success rate, but also the failure modes most likely to arise in practical use.

@@

\subsection{Problematic Cases}
\label{subsec:problematic-cases}

Table~\ref{tab:problematic_cases} presents examples of problematic cases in which the system was asked to generate a valid
Fluid expression.
In these cases, producing a valid expression is challenging because the input data (or the NL query) contains ambiguities
that prevent the construction of a well-formed expression.

For instance, in the \textit{missing-data} problem, the \InterpretationAgent should generate a function to compute
 the average of the emissions. However, the input data does not provide information for all months.

In such cases, the system was unable to generate a general expression and instead produced the expected value directly.

>>>>

We investigated the following research questions, which probe the limits of the \InterpretationAgent's ability
to reliably provide a data-driven interpretation of natural language. By contrasting genuine uncertainty with
misleading or ill-specified inputs, we aim to understand not only the model's raw success rate, but also the
failure modes most likely to arise in practical use.

\paragraph{RQ1. How robust is the system when provided with incomplete or ambiguous information?}

In real-world authoring tasks, text fragments may omit key criteria (e.g. \textbf{?}) or contain vague
identifiers (e.g. \textbf{?} without further context). We therefore ask how performance degrades when the
specification is underspecified or when multiple interpretations are possible. Here, we distinguish between
two forms of ambiguity: (1) domain-level ambiguity, which reflects genuine uncertainty in the data (e.g. two
models achieving identical scores). (2) specification-level ambiguity, which arises from incomplete or
imprecise descriptions in the query (e.g. missing fields that would otherwise resolve the expression
uniquely).

\paragraph{RQ2. How sensitive is the system to misleading or adversarial information?}

Authors may (intentionally or unintentionally) introduce misleading elements into the text. These include
inconsistent use of language, incorrect references to numerical values, or contradictory claims. We ask how
the system responds in such cases: does it fail gracefully, does it default to literal reproduction of the
misleading fragment, or does it attempt to generate a plausible, but incorrect, expression?
Table~\ref{tab:problematic_cases} shows some of cases which are ``problematic'' in this sense; in these cases,
producing a valid expression is likely to be challenging because of ambiguities in the input data or
accompanying natural language.

@@
